---
title: "googleAnalyticsModelR"
date: 11th Feb, 2019
output: md_document
---

# googleAnalyticsModelR

Creating ready made models to work with `googleAnalyticsR` data

## Setup

```r
install.packages(c("remotes","googleAnalyticsR"))
remotes::install_github("IronistM/googleAnalyticsModelR")
```

## Useage

For end users, they can just load the model then apply it to their data:

```{r message=FALSE, warning=FALSE}
library(googleAnalyticsR)  # assume auto-authentication
library(googleAnalyticsModelR)

# fetches data and outputs decomposition
my_viewid <- 81416156
decomp_ga <- "inst/models/decomp_ga.gamr"
d1 <- ga_model(my_viewid, model = decomp_ga)

#repeat with another viewId
d2 <- ga_model(123875646, model = decomp_ga)

# Example CausalImpact
ci <- ga_model(81416156, model = "inst/models/causalImpact_model.gamr", 
         event_date = Sys.Date() - 51, 
         predictors = "Direct", 
         response = "Organic Search")

```

## Creating model `.gamr` objects

To create your own models, you need to predefine all the functions to look after the fetching, modelling and viewing of the data.  You then pass those functions to the `ga_model_make()` function.

The functions need to follow these specifications:

* `data_f` - A function to collect the data you will need. The first argument should be the `view_id` which will be pass the viewId of Google Analytics property to fetch data from.
* `model_f` - A function to work with the data you have fetched. The first argument should be the data.frame that is produced by the data fetching function, `data_f()`.
* `output_f` - A function to plot the data. The first argument should be the data.frame that is produced by the model function, `model_f()`. 
* All functions you create must include `...` as an argument. 
* All functions must use different arguments (apart from `...`), to avoid clashes.

If you want to also create the Shiny modules, then you also need to specify:
* `outputShiny` - the output function for the UI, such as `plotOutput`
* `renderShiny` - the render function for the server, such as `renderPlot`

You then supply supporting information to make sure the user can run the model:

* `required_columns` - Specification of which columns the data will fetch.  It will fail if they are not present.
* `required_packages` - The packages the end user needs to have installed to run your functions.
* `description` - A sentence on what the model is so they can be distinguished. 

To create the example model above, the above was applied as shown below:

```r
get_model_data <- function(viewId,
                           date_range = c(Sys.Date()- 300, Sys.Date()),
                           ...){
   google_analytics(viewId,
                    date_range = date_range,
                    metrics = "sessions",
                    dimensions = "date",
                    max = -1)
 }

decompose_sessions <- function(df, ...){
   decompose(ts(df$sessions, frequency = 7))
 }
 
decomp_ga <- ga_model_make(get_model_data,
                           required_columns = c("date", "sessions"),
                           model_f = decompose_sessions,
                           output_f = graphics::plot,
                           description = "Performs decomposition and creates a plot",
                           outputShiny = shiny::plotOutput,
                           renderShiny = shiny::renderPlot)
```

## Advanced use

The more arguments you provide to the model creation functions, the more complicated it is for the end user, but the more flexible the model.  It is suggested making several narrow useage models is better than one complicated one.

For instance, you could modify the above model to allow the end user to specify the metric, timespan and seasonality of the decomposition:

```{r}
get_model_data <- function(viewId,
                           date_range = c(Sys.Date()- 300, Sys.Date()),
                           metric,
                           ...){
   o <- google_analytics(viewId,
                    date_range = date_range,
                    metrics = metric,
                    dimensions = "date",
                    max = -1)
    # rename the metric column so its found for modelling
    o$the_metric <- o[, metric]
    
    o
    
 }

decompose_sessions <- function(df, frequency, ...){
   decompose(ts(df$the_metric, frequency = frequency))
 }
 
decomp_ga_advanced <- ga_model_make(get_model_data,
                           required_columns = c("date"), # less restriction on column
                           model_f = decompose_sessions,
                           output_f = graphics::plot,
                           description = "Performs decomposition and creates a plot",
                           outputShiny = shiny::plotOutput,
                           renderShiny = shiny::renderPlot)
                           
```

It would then be used via:

```{r message=FALSE, warning=FALSE}
result <- ga_model(81416156, decomp_ga_advanced, metric="users", frequency = 30)
str(result, max.level = 1)
```

### Working with the model object

The model objects prints to console in a friendly manner:

```{r}
decomp_ga_advanced
```
You can save and load model objects from a file.  It is suggested to save them with the `.gamr` suffix.

```r
# save model to a file
ga_model_save(decomp_ga_advanced, filename = "my_model.gamr")

# load model again
ga_model_load("my_model.gamr")
```

You can use models directly from the file:

```r
ga_model(81416156, "my_model.gamr")
```

If you need to change parts of a model, `ga_model_edit()` lets you change individual aspects:

```{r}
ga_model_edit(decomp_ga_advanced, description = "New description")
```

You can also pass it the filename, which will load, make the edit, then save the model to disk again:

```r
ga_model_edit("my_model.gamr", description = "New description")
```

## More complicated example

## CausalImpact example

To make your own portable GA Effect, this model uses the CausalImpact and dygraphs libraries to make a plot of your GA data.

This example model is available via `ga_model_example("ga-effect.gamr")`

```{r}
library(googleAnalyticsR)

get_ci_data <- function(viewId, 
                        date_range = c(Sys.Date()-600, Sys.Date()),
                        ...){
  
  google_analytics(viewId, 
                   date_range = date_range,
                   metrics = "sessions",
                   dimensions = c("date", "channelGrouping"), 
                   max = -1)
}

# response_dim is the channel to predict.
# predictors help with forecast
do_ci <- function(df, 
                  event_date,
                  response = "Organic Search",
                  predictors = c("Video","Social","Direct"),
                  ...){
  
  message("CausalImpact input data columns: ", paste(names(df), collapse = " "))
  # restrict to one response 
  stopifnot(is.character(response), 
            length(response) == 1,
            assertthat::is.date(event_date),
            is.character(predictors))
  
  pivoted <- df %>% 
    tidyr::spread(channelGrouping, sessions)
  
  stopifnot(response %in% names(pivoted))
  
  ## create a time-series zoo object
  web_data_xts <- xts::xts(pivoted[-1], order.by = as.Date(pivoted$date), frequency = 7)
  
  pre.period <- as.Date(c(min(df$date), event_date))
  post.period <- as.Date(c(event_date + 1, max(df$date)))
  
  predictors <- intersect(predictors, names(web_data_xts))

  ## data in order of response, predictor1, predictor2, etc.
  model_data <- web_data_xts[,c(response,predictors)]
  
  # deal with names
  names(model_data) <- make.names(names(model_data))
  # remove any NAs
  model_data[is.na(model_data)] <- 0

  CausalImpact::CausalImpact(model_data,  pre.period, post.period)

}

dygraph_plot <- function(impact, event_date, ...){
  require(dygraphs)
  ## the data for the plot is in here
  ci <- impact$series
  
  ci <- xts::xts(ci)

  ## the dygraph output
  dygraph(data=ci[,c('response', 'point.pred', 'point.pred.lower', 'point.pred.upper')], 
          main="Expected (95% confidence level) vs Observed", group="ci") %>%
    dyEvent(x = event_date, "Event") %>%
    dySeries(c('point.pred.lower', 'point.pred','point.pred.upper'), 
             label='Expected') %>%
    dySeries('response', label="Observed")
}

req_packs <- c("CausalImpact", "xts", "tidyr", "googleAnalyticsR", "assertthat", "dygraphs")

ci_model <- ga_model_make(get_ci_data,
                          required_columns = c("date","channelGrouping","sessions"),
                          model_f = do_ci,
                          output_f = dygraph_plot,
                          required_packages = req_packs,
                          description = "Causal Impact on channelGrouping data",
                          outputShiny = dygraphs::dygraphOutput,
                          renderShiny = dygraphs::renderDygraph)
# print out model details
ci_model
```


```r
# save it to a file for use later
ga_model_save(ci_model, "causalImpact_model.gamr")
```

To use:

```{r message=FALSE, warning=FALSE}
library(googleAnalyticsR)
library(xts)
library(tidyr)
library(dygraphs)

ci <- ga_model(81416156, ci_model, event_date = as.Date("2019-01-01"))
```

Similarly, you can launch this in a Shiny app by slightly modifying the example above.  

This is available within the package via `shiny::runApp(system.file("shiny/models-ga-effect", package="googleAnalyticsR"))`

### Using model objects within functions

You can go more meta by encasing the model definition and use in another function.  This is used by this example of [Dartistic's example "Time normalised pageviews"](http://www.dartistics.com/googleanalytics/int-time-normalized.html) by Tim Wilson.

To use the end result:

```{r message=FALSE, warning=FALSE}
library(googleAnalyticsR)
library(googleAnalyticsModelR)

output <- ga_time_normalised(81416156)

```

`ga_time_normalised()` wraps a call to `ga_model()`;

```{r}
#' Time normalised traffic
#'
#' Based on \url{http://www.dartistics.com/googleanalytics/int-time-normalized.html} by Tim Wilson
#'
#' @param viewId The viewId to use
#' @param first_day_pageviews_min threshold for first day of content
#' @param total_unique_pageviews_cutoff threshold of minimum unique pageviews
#' @param days_live_range How many days to show
#' @param page_filter_regex Select which pages to appear
#'
#' @return A \link[googleAnalyticsR]{ga_model} object
#'
#' @export
ga_time_normalised <- function(viewId,
                               first_day_pageviews_min = 2,
                               total_unique_pageviews_cutoff = 500,
                               days_live_range = 60,
                               page_filter_regex = ".*"){

  model <- ga_model_load(filename = "inst/models/time-normalised.gamr")

  ga_model(viewId,
           model,
           first_day_pageviews_min = first_day_pageviews_min,
           total_unique_pageviews_cutoff = total_unique_pageviews_cutoff,
           days_live_range = days_live_range,
           page_filter_regex = page_filter_regex)

}
```

The model itself is created by issuing `make_time_normalised()` and wraps the code ported from the Dartistics code example, putting it in the right function formats:

```{r}
#' Run this manually when you want to alter the saved model
#' @noRd
make_time_normalised <- function(){

  data_f <- function(viewId, page_filter_regex, ...){
    page_filter_object <- dim_filter("pagePath",
                                     operator = "REGEXP",
                                     expressions = page_filter_regex)
    page_filter <- filter_clause_ga4(list(page_filter_object),
                                     operator = "AND")

    google_analytics(viewId = viewId,
                     date_range = c(Sys.Date() - 365, Sys.Date() - 1),
                     metrics = "uniquePageviews",
                     dimensions = c("date","pagePath"),
                     dim_filters = page_filter,
                     anti_sample = TRUE)

  }
  model_f <- function(ga_data,
                      first_day_pageviews_min,
                      total_unique_pageviews_cutoff,
                      days_live_range,
                      ...){
    normalize_date_start <- function(page){
      ga_data_single_page <- ga_data %>% filter(pagePath == page)
      first_live_row <- min(which(ga_data_single_page$uniquePageviews > first_day_pageviews_min))
      ga_data_single_page <- ga_data_single_page[first_live_row:nrow(ga_data_single_page),]
      normalized_results <- data.frame(date = seq.Date(from = min(ga_data_single_page$date),
                                                       to = max(ga_data_single_page$date),
                                                       by = "day"),
                                       days_live = seq(min(ga_data_single_page$date):
                                                         max(ga_data_single_page$date)),
                                       page = page) %>%
        left_join(ga_data_single_page) %>%
        mutate(uniquePageviews = ifelse(is.na(uniquePageviews), 0, uniquePageviews)) %>%
        mutate(cumulative_uniquePageviews = cumsum(uniquePageviews)) %>%
        select(page, days_live, uniquePageviews, cumulative_uniquePageviews)
    }

    pages_list <- ga_data %>%
      group_by(pagePath) %>% summarise(total_traffic = sum(uniquePageviews)) %>%
      filter(total_traffic > total_unique_pageviews_cutoff)

    ga_data_normalized <- map_dfr(pages_list$pagePath, normalize_date_start)

    ga_data_normalized %>% filter(days_live <= days_live_range)
  }

  output_f <- function(ga_data_normalized, ...){
    gg <- ggplot(ga_data_normalized,
                 mapping=aes(x = days_live, y = cumulative_uniquePageviews, color=page)) +
      geom_line() +                                          # The main "plot" operation
      scale_y_continuous(labels=comma) +                     # Include commas in the y-axis numbers
      labs(title = "Unique Pageviews by Day from Launch",
           x = "# of Days Since Page Launched",
           y = "Cumulative Unique Pageviews") +
      theme_light() +                                        # Clean up the visualization a bit
      theme(panel.grid = element_blank(),
            panel.border = element_blank(),
            legend.position = "none",
            panel.grid.major.y = element_line(color = "gray80"),
            axis.ticks = element_blank())

    ggplotly(gg)
  }
  required_columns <- c("date","pagePath","uniquePageviews")
  required_packages <- c("plotly", "scales", "dplyr", "purrr", "ggplot2")

  model <- ga_model_make(
    data_f = data_f,
    required_columns = required_columns,
    model_f = model_f,
    output_f = output_f,
    required_packages = required_packages,
    description = "Cumalitive visualisation of time-normalised traffic",
    outputShiny = plotly::plotlyOutput,
    renderShiny = plotly::renderPlotly
  )

  ga_model_save(model, filename = "inst/models/time-normalised.gamr")

  model

}
```

Say we now save this model to `"time-normalised.gamr"`.  

You can use the module functions to turn it into a Shiny app:

In this case, we need to build the UI, and the input selections:

```r
library(shiny)             # R webapps
library(gentelellaShiny)   # ui theme
library(googleAuthR)       # auth login
library(googleAnalyticsR) # get google analytics

# the libraries needed by the model
library(dplyr)
library(plotly)
library(scales)
library(ggplot2)
library(purrr)

# set your GCP project for the auth
gar_set_client(web_json = "your-client-web.json",
               scopes = "https://www.googleapis.com/auth/analytics.readonly")

model <- ga_model_load(filename = "time-normalised.gamr")

ui <- gentelellaPage(
  menuItems = list(sideBarElement(googleAuthUI("auth_menu"))),
  title_tag = "GA Time Normalised Pages",
  site_title = a(class="site_title", icon("phone"), span("Time normalised")),
  footer = "Made in Denmark",

  # shiny UI elements
  column(width = 12, authDropdownUI("auth_dropdown", inColumns = TRUE)),
  numericInput("first_day", "First day minimum pageviews", value = 2, min=0, max=100),
  numericInput("total_min_cutoff", "Minimum Total pageviews", value = 500, min = 0, max = 1000),
  numericInput("days_live", label = "Days Live", value = 60, min = 10, max = 400),
  textInput("page_regex", label = "Page filter regex", value = ".*"),
  h3("Time Normalised pages"),
  model$shiny_module$ui("model1"),
  br()

)

server <- function(input, output, session) {

  gar_shiny_auth(session)

  al <- reactive(ga_account_list())

  # module for authentication
  view_id <- callModule(authDropdown, "auth_dropdown", ga.table = al)

  callModule(model$shiny_module$server,
             "model1",
             view_id = view_id,
             first_day_pageviews_min = reactive(input$first_day),
             total_unique_pageviews_cutoff = reactive(input$total_min_cutoff),
             days_live_range = reactive(input$days_live),
             page_filter_regex = reactive(input$page_regex))

}
# Run the application
shinyApp(gar_shiny_ui(ui, login_ui = silent_auth), server)
```
